\subsection{Response format}

Return the result strictly in two sections. Text outside the sections is not allowed.

\subsection{Section 1: Mermaid}

Format: \texttt{flowchart LR}.

Tables are trapezoids with the \texttt{tbl\_} prefix:

\begin{verbatim}
tbl_user[/"User (entity)"/]
\end{verbatim}

Functions are rounded rectangles with the \texttt{fn\_} prefix:

\begin{verbatim}
fn_register_user(["RegisterUser"])
\end{verbatim}

Allowed edge types (must be specified in the edge label):

\begin{itemize}
\item \texttt{consumes}: table $\to$ function
\item \texttt{produces}: function $\to$ table
\item \texttt{reads}: table $\to$ function
\item \texttt{writes}: function $\to$ table
\item \texttt{triggers}: function $\to$ function
\end{itemize}

Forbidden: \texttt{table $\to$ table}, \texttt{function $\to$ function} without \texttt{triggers}.

\subsection{Section 2: JSON}

A single object with two keys: \texttt{functions} and \texttt{tables}.

\subsubsection{\texttt{functions}}

The key is the function name (\texttt{PascalCase}, verb + noun). The value is an object:

\begin{lstlisting}
{
  "purpose": "one sentence",
  "processing": ["step 1", "step 2", "..."],
  "input": {
    "consumes": { "TableName": "short description" },
    "reads":    { "TableName": "short description" }
  },
  "output": {
    "produces": { "TableName": "short description" },
    "writes":   { "TableName": "short description" }
  },
  "errors": [
    { "condition": "condition", "result": "result" }
  ],
  "triggers": [
    { "function": "FunctionName", "condition": "condition" }
  ],
  "contract": {
    "Cin":  ["tables"],
    "Cout": ["tables"],
    "R":    ["tables"],
    "W":    ["tables"],
    "Tin":  ["functions"],
    "Tout": ["functions"]
  }
}
\end{lstlisting}

Rules:

\begin{itemize}
\item \texttt{consumes} / \texttt{reads} / \texttt{produces} / \texttt{writes}: the value must be a short string only, not an object.
\item \texttt{processing}: each element must describe one action on one line.
\item \texttt{contract}: arrays are treated as sets; order is irrelevant and duplicates are forbidden.
\item Empty fields: \texttt{\{\}} for objects, \texttt{[]} for arrays.
\end{itemize}

\subsubsection{\texttt{tables}}

The key is the table name (\texttt{PascalCase}, singular). The value is an object:

\begin{lstlisting}
{
  "x-kind": "entity | event | reference | log | projection | error",
  "x-pk": "field_name",
  "properties": {
    "field_name": { "type": "type", "format": "format" }
  },
  "required": ["required fields"]
}
\end{lstlisting}

Field-level extensions (inside \texttt{properties}):

\begin{itemize}
\item \texttt{"x-fk": "Table.field"} --- foreign key.
\item \texttt{"x-unique": true} --- unique value.
\end{itemize}

\texttt{x-pk} and \texttt{x-fk} are \texttt{single-field} only.

\vspace{0.5em}
\noindent\textbf{Field types.}
\vspace{0.5em}

\begin{tabular}{lll}
\hline
\textbf{CGD type} & \textbf{JSON Schema type} & \textbf{Note} \\
\hline
\texttt{text} & \texttt{string} & \\
\texttt{number} & \texttt{number} & \texttt{integer} is allowed \\
\texttt{bool} & \texttt{boolean} & \\
\texttt{date} & \texttt{string} + \texttt{format: date-time} & \\
\texttt{enum} & \texttt{string} + \texttt{enum: [values]} & \\
\texttt{ref} & target PK type + \texttt{x-fk} & \\
\hline
\end{tabular}

\vspace{0.5em}
Field names and enum values: \texttt{snake\_case}.

\subsection{Section priority}

If Mermaid and JSON diverge, JSON is the source of truth.

\subsection{Consistency}

The sections must be consistent:

\begin{itemize}
\item Every node in Mermaid = an entry in \texttt{functions} or \texttt{tables}.
\item Every edge in Mermaid = an element of \texttt{contract}.
\item Every element of \texttt{contract} = an edge in Mermaid.
\item \texttt{kind} in a Mermaid table label = \texttt{x-kind} in \texttt{tables}.
\item Keys in \texttt{consumes} / \texttt{reads} / \texttt{produces} / \texttt{writes} = the corresponding arrays in \texttt{contract}.
\item Values in \texttt{triggers[*].function} = the \texttt{Tout} array in \texttt{contract}.
\end{itemize}
